High-level cognition during story listening is reflected in high-order dynamic correlations in neural activity patterns

### 2020-11-11 paper/main.tex
finished addressing reviewer comments

We used a subset of the story listening and rest data to train
across-participant classifiers to decode listening times using a blend
of neural features (comprising neural activity patterns, as well as
different orders of dynamic correlations between those patterns that
were inferred using our computational framework).  We found that both
the PCA-based and eigenvector centrality-based approaches yielded
neural patterns that could be used to decode accurately (i.e., well
above chance).  Both approaches also yielded the best decoding
accuracy for data collected during (intact) story listening when
high-order (PCA: second-order; eigenvector centrality: fourth-order)
dynamic correlation patterns were included as features.  When we
trained classifiers on the scrambled stories or resting state data,
only (relatively) lower-order dynamic patterns were informative to the decoders.
Taken together, our results indicate that high-level cognition is
supported by high-order dynamic patterns of communication between
brain structures.

@@

The notion that cognition is reflected in (and possibly mediated by)
patterns of first-order network dynamics has been suggested by or
proposed in myriad empirical studies and
reviews~\citep[e.g.,][]{DemeEtal19, Turk13, LuriEtal18, FongEtal19,
  ParkEtal18b, PretEtal17, RoyEtal19, LiegEtal19, ZouEtal19,
  ChanGlov10, GonzEtal19}.  Our study extends this line of work by
finding cognitively relevant \textit{higher-order} network dynamics
that reflect ongoing cognition.  Our findings complement other work
that uses graph theory and topology to characterize how brain networks
reconfigure during cognition~\citep[e.g.,][]{BassEtal06, ZhenEtal19,
  McInJirs19, TokeSomm19, SizeEtal18, ReimEtal17, BetzEtal19}.

@@

Identifying high-order network dynamics associated with high-level
cognition required several important methods advances.  First, we used
kernel-based dynamic correlations to extended the notion of (static)
inter-subject functional connectivity~\citep{SimoEtal16} to a dynamic
measure of inter-subject functional connectivity (DISFC) that does not
rely on sliding windows, and that may be computed at individual
timepoints.  This allowed us to precisely characterize stimulus-evoked
network dynamics that were similar across individuals.  Second, we
developed a computational framework for efficiently and scalably estimating
high-order dynamic correlations.  Our approach uses dimensionality
reduction algorithms and graph measures to obtain low-dimensional
embeddings of patterns of network dynamics.  Third, we developed an
analysis framework for identifying robust decoding results by carrying
out our analyses using a range of parameter values and then
identifying which results were robust to specific parameter choices.

\subsection*{Concluding remarks}
The complex hierarchy of dynamic interactions that underlie our
thoughts is perhaps the greatest mystery in modern science.  Methods
for characterizing the dynamics of high-order correlations in neural
data provide a window into the neural basis of cognition.  By showing
that high-level cognition is reflected in high-order network dynamics,
we have elucidated the next step on the path towards understanding the
neural basis of cognition.

>>>>

We used a subset of the story listening and rest data to train
across-participant classifiers to decode listening times (of groups of
participants) using a blend of neural features (comprising neural
activity patterns, as well as different orders of dynamic correlations
between those patterns that were inferred using our computational
framework).  We found that both the PCA-based and eigenvector
centrality-based approaches yielded neural patterns that could be used
to decode accurately (i.e., well above chance).  Both approaches also
yielded the best decoding accuracy for data collected during (intact)
story listening when high-order (PCA: second-order; eigenvector
centrality: fourth-order) dynamic correlation patterns were included
as features.  When we trained classifiers on the scrambled stories or
resting state data, only (relatively) lower-order dynamic patterns
were informative to the decoders.  Taken together, our results
indicate that high-level cognition is supported by high-order dynamic
patterns of communication between brain structures.

@@

The notion that cognition is reflected in (and possibly mediated by)
patterns of first-order network dynamics has been suggested by or
proposed in myriad empirical studies and
reviews~\citep[e.g.,][]{DemeEtal19, Turk13, LuriEtal18, FongEtal19,
  ParkEtal18b, PretEtal17, MannEtal18, RoyEtal19, LiegEtal19, ZouEtal19,
  ChanGlov10, GonzEtal19}.  Our study extends this line of work by
finding cognitively relevant \textit{higher-order} network dynamics
that reflect ongoing cognition.  Our findings complement other work
that uses graph theory and topology to characterize how brain networks
reconfigure during cognition~\citep[e.g.,][]{BassEtal06, ZhenEtal19,
  McInJirs19, TokeSomm19, SizeEtal18, ReimEtal17, BetzEtal19}.

@@

Identifying high-order network dynamics associated with high-level
cognition required several important methods advances.  First, we used
kernel-based dynamic correlations to extended the notion of (static)
inter-subject functional connectivity~\citep{SimoEtal16} to a dynamic
measure of inter-subject functional connectivity (DISFC) that does not
rely on sliding windows~\citep[e.g., as in][]{MannEtal18}, and that
may be computed at individual timepoints.  This allowed us to
precisely characterize stimulus-evoked network dynamics that were
similar across individuals.  Second, we developed a computational
framework for efficiently and scalably estimating high-order dynamic
correlations.  Our approach uses dimensionality reduction algorithms
and graph measures to obtain low-dimensional embeddings of patterns of
network dynamics.  Third, we developed an analysis framework for
identifying robust decoding results by carrying out our analyses using
a range of parameter values and then identifying which results were
robust to specific parameter choices.

\subsection*{Concluding remarks}
Methods for characterizing the dynamics of high-order correlations in
neural data provide a window into the neural basis of cognition.  By
showing that high-level cognition is reflected in high-order network
dynamics, we have elucidated the next step on the path towards
understanding the neural basis of cognition.
